---
title: "Project Part 1"
author: "Kris Stewart"
date: "2025-02-02"
output: word_document
---

Load in all the relevant Libraries. 

```{r, message=FALSE}

library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)
library(dplyr)
library(GGally)
library(caret)
library(ranger)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(naniar) #visualizing missingness
library(skimr) #alternative way to view dataset summaries
library(UpSetR) #visualizing missingness
library(dplyr)
library(glmnet) #for Lasso, ridge, and elastic net models 
library(ggcorrplot) #create an alternative to ggcorr plots
library(MASS) #access to forward and backward selection algorithms
library(leaps) #best subset selection
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting
```

Read in the data sets.

```{r, message=FALSE}

sample_submission <- read_csv("~/Documents/School/Predictive Analytics/Project/Part 1/sample_submission.csv")
test <- read_csv("~/Documents/School/Predictive Analytics/Project/Part 1/test.csv")
train <- read_csv("~/Documents/School/Predictive Analytics/Project/Part 1/train.csv")

```

Evaluate the data set given

```{r}

str(train)
summary(train)
skim(train)

```
Transform data type to factors. 

```{r}

train <- train %>% mutate(product_code = as_factor(product_code)) %>%
  mutate(attribute_0 = as_factor(attribute_0)) %>% 
  mutate(attribute_1 = as_factor(attribute_1)) %>% 
  mutate(attribute_2 = as_factor(attribute_2)) %>% 
  mutate(attribute_3 = as_factor(attribute_3)) %>%
  mutate(failure = as_factor(failure))

```

Evaluate the Missingness of data

```{r warning=FALSE}
gg_miss_var(train)
gg_miss_case(train) 
vis_miss(train )
```


Basic plots of variables to understand distribution

```{r }

ggplot(data = train, aes(loading, fill=failure)) + geom_histogram()
ggplot(train, aes(x=product_code, fill = failure)) + geom_bar() + theme_bw()
ggplot(train, aes(x=attribute_0, fill = failure)) + geom_bar() + theme_bw()
ggplot(train, aes(x=attribute_1, fill = failure)) + geom_bar() + theme_bw()
ggplot(train, aes(x=attribute_2, fill = failure)) + geom_bar() + theme_bw()
ggplot(train, aes(x=attribute_3, fill = failure)) + geom_bar() + theme_bw()


```

Create a new visual to show which factors of attributes have highest failure rate

```{r message=FALSE}
train %>%
  dplyr::select(where(is.factor), failure) %>%
  pivot_longer(-failure, names_to = "Variable", values_to = "Value") %>%
  group_by(Variable, Value) %>%
  summarize(FailureRate = mean(failure == "Yes") * 100) %>%
  ggplot(aes(x = reorder(Value, FailureRate), y = FailureRate, fill = Variable)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(~Variable, scales = "free_y") +
  labs(title = "Failure Rate by Factor Variables", y = "Failure Rate (%)", x = "") +
  theme_minimal()
```


Correlation of variables

```{r warning=FALSE}

ggcorr(train)

```

Check the importance of the variables.

```{r}

rf_model <- ranger(failure ~ ., data = train, importance = "impurity")
importance <- as.data.frame(rf_model$variable.importance) %>%
  rownames_to_column(var = "Feature") %>%
  arrange(desc(rf_model$variable.importance))

# Plot Feature Importance
ggplot(importance, aes(x = reorder(Feature, rf_model$variable.importance), y = rf_model$variable.importance, fill = Feature)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Feature Importance in Predicting Failure", x = "Feature", y = "Importance") +
  theme_minimal()

# Print the most important variables
print(importance)
```


Check the initial p-values of the data to see any initial significance.

```{r}

model <- glm(failure ~., data = train, family = binomial())
summary(model)

```



# STOP


```{r}

ggcorr(train)
ggpairs(train, columns = c("failure","measurement_0", "measurement_1","measurement_2", "measurement_3", "measurement_4","measurement_5"))
ggpairs(train, columns = c("failure","measurement_6", "measurement_7","measurement_8", "measurement_9", "measurement_10","measurement_11"))
ggpairs(train, columns = c("failure","measurement_12", "measurement_13","measurement_14", "measurement_15", "measurement_16"))
ggpairs(train, columns = c("failure","product_code", "attribute_0", "attribute_1", "attribute_2", "attribute_3"))

```
```{r}
#train = train %>% dplyr::select(1:16,26) 

#skim(train)
```


```{r}

#imputed_train <- mice(train, m=5, method = "pmm", seed =123)
#train <- complete(imputed_train)

#str(train)
#summary(train)
#skim(train)
```

```{r}
train_model = 
  logistic_reg() %>% 
  set_engine("glm") 

train_recipe = recipe(failure ~., train) %>%
  step_other(product_code, threshold = 0.01) %>% 
  step_other(attribute_0, threshold = 0.01) %>% 
  step_dummy(all_nominal(), -all_outcomes()) 


logreg_wf = workflow() %>%
  add_recipe(train_recipe) %>% 
  add_model(train_model)

train_fit = fit(logreg_wf, train)
```

```{r}
summary(train_fit$fit$fit$fit)
```






```{r}

set.seed(12345)
train_split = initial_split(train, prop = 0.70, strata = failure)
train1 = training(train_split)
test1 = testing(train_split)

```


```{r}

levels(train1$failure)

```


```{r}

train_model = 
  logistic_reg(mode = "classification") %>% 
  set_engine("glm") 

train_recipe = recipe(failure ~ product_code + attribute_0 + attribute_1 + attribute_2 + attribute_3, train1) %>%
  step_dummy(all_nominal(), -all_outcomes()) 

logreg_wf = workflow() %>%
  add_recipe(train_recipe) %>% 
  add_model(train_model)

train_fit = fit(logreg_wf, train1)

```


```{r}
summary(train_fit$fit$fit$fit)
```








